---
title: "607 Homework 1"
author: "Victoria McEleney"
date: "`r Sys.Date()`""
output: github_document
---
#Candy Power Ranking

To figure out what Halloween candy people prefer dozens of fun-sized candy varieties were pitted against one another.  8,371 different IP addresses voted on about 269,000 randomly generated matchups.

https://fivethirtyeight.com/videos/the-ultimate-halloween-candy-power-ranking/

#Load libraries.

```{r load-packages, message=FALSE}
library(tidyverse)
```

#Load the .csv into a data.frame

```{r}
candy <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv", sep=",")
candy
```

#Subset & sort winpercent in descending order
The candy with highest 12 winpercent had chocolate.

```{r}
candy %>% 
  select(competitorname, chocolate, winpercent) %>% 
  arrange(desc(winpercent))
```

#Plot colored by chocolate

```{r}
ggplot(candy) +
geom_point(mapping = aes(x = competitorname, y = winpercent, color = chocolate)) +
labs(title = "Chocolate vs. Non-chocolate") +
theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

