---
title: "Week3 - Character Manipulation and Date Processing"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Tibble operations
library(tidyverse)

# For data cleanup
library(stringr)

```

#### 1. & 2. Answered by group member Cliff Lee: 

<https://github.com/cliftonleesps/607_acq_mgt/blob/main/week3/homework3.Rmd>

#### 3. Describe, in words, what these expressions will match:


##### (.)\1\1 

will match any 3 consecutive characters (ex: CCC)


##### "(.)(.)\\2\\1" 

will match any 2 characters followed by the same 2 captured characters in reverse order (ex: eppe)


##### (..)\1 

will match any 2 characters followed by the same 2 characters (ex: anan)


##### "(.).\\1.\\1" 

will match any character followed by any character, followed by the 1st captured character, followed by any character apart from 1st captured character, followed by the 1st captured character (ex: anana, anama)


##### "(.)(.)(.).*\\3\\2\\1" 

will match any 3 characters, followed by any or no characters, followed by the first captured 3 characters in reverse order (ex: aprrpa, apricotrpa)



#### 4. Construct regular expressions to match words that:

##### Start and end with the same character.


```{r}

words[str_detect(words, "^(.)((.*\\1$)|\\1?$)")]
```


##### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)


```{r}
str_view(words, "([a-z][a-z]).*\\1", match = TRUE)

```


##### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


```{r}
str_view(words, "([a-z]).*\\1.*\\1", match = TRUE)

```

