---
title: "607 HW 7"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more than one author. For each book, include the title, authors, and two or three other attributes that you find interesting. Take the information that you’ve selected about these three books, and separately create three files which store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”, “books.xml”, and “books.json”). Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames. Are the three data frames identical?

## Load libraries
```{r}
library(XML)
library(xml2)
library(jsonlite)
library(arsenal)
```

## Create xml, upload to Github, and read xml file using read_xml from xml2 package
```{r}
booksx = read_xml ("https://raw.githubusercontent.com/vmceleney/DATA-607/main/Books.xml")
```

## Parse xml using xmlParse functon from XML package
```{r}
booksxp <- xmlParse(booksx)
booksxp
```

## Get the top-level node and its elements from the parsed XMLInternalDocument using xmlRoot from XML package
```{r}
booksxroot <- xmlRoot(booksxp)
booksxroot
```

## Extract data from XMLInternalElementNode using xmltoDataFrame from the XML package
```{r}
booksxd <- xmlToDataFrame(booksxroot)
booksxd
```

## create html, upload to Github, and read .html file using ready_html from xml2 package
```{r}
booksh = read_html("https://raw.githubusercontent.com/vmceleney/DATA-607/main/books.html")
booksh
```
## Parse html using htmlParsen function from XML package
```{r}
bookshp <- htmlParse(booksh)
```

## Examine nodes and children
```{r}
xmlRoot(bookshp)
xmlChildren(bookshp)

```

## Read data using HTMLTable from XML package
```{r}
booksht <- readHTMLTable(bookshp, which = 1)
booksht
```

## create json, upload to Github, and convert .json file to R objects using  from jsonlite package
```{r}
booksj = fromJSON("https://raw.githubusercontent.com/vmceleney/DATA-607/main/Books.json")
booksj
```
# How do the dataframes created from xml, html, & json compare?

## Compare dataframes using comparedf from arsenal package
```{r}
comparedf(booksxd, booksht)
comparedf(booksxd, booksj)
comparedf(booksht, booksj)
```
# Conclusion: The 3 dataframes are identical.
