---
title: "607 HW 9"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
Sign up for an API key and choose one of the New York Times APIs. construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

## Load libraries
```{r}
library(jsonlite)
library(dplyr)
```

## Access NYTimes Movie API and query by keywords Kate Beckinsale 
```{r}
df <- fromJSON("https://api.nytimes.com/svc/movies/v2/reviews/search.json?query=kate+beckinsale&api-key=t8jjTBB9jh07Izw30VZciBZVqvsFQUV7", flatten = TRUE)
df <- data.frame(df$results)
```

## Conclusion
We have returned the maxiumum 20 rows allowed by accessing the NYTimes Movie API and querying by keywords Kate Beckinsale using the jsonlite library. The display_title column is consistent with movies starring Kate Beckinsale as listed on imdb.com.

