---
title: "607 Project 2 - Part 2"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Practice preparing different datasets for downstream analysis work. Choose any three of the “wide” datasets identified in the Week 5 Discussion items. For each of the three chosen datasets:

(1) Create a .CSV file that includes all of the information included in the dataset. 

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

(3) Perform the analysis requested in the discussion item.

### This table contains the running times for four runners, taken immediately before starting a new training regime and then for three months thereafter.  Visualize the progress of these runners after tidying the data. Find who improved the most from the training program.

## Load libraries

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(tidyverse)
```

## Store data in Excel, save as .csv, and read .csv file into R:
```{r}
running_times <- read_csv("https://raw.githubusercontent.com/vmceleney/DATA-607/main/running_times.csv", 
    skip = 2)
running_times
```

# Rename columns:
```{r}
running_times <- rename(running_times, `0` = `Month 0`, `1` = `Month 1` , `2` = `Month 2`, `3` = `Month 3`)
```

## Which runner's time improved the most?
## calculate improvement rate:
```{r}
running_times <- running_times %>%  
  mutate(improvement_rate = 1 - (`3` / `0`))
running_times
```
### Joana's run time decreased / improved the most at 4.08%

## Reshape data longer:
```{r}
running_times_longer <- pivot_longer(running_times, cols = 2:5, names_to ="Month",
values_to = "times")
running_times_longer
```

## Convert Month to numeric:
```{r}
running_times_longer$Month <- as.numeric(running_times_longer$Month)
```

## What else can we find out by visualizing data?
## Plot data:
```{r}
ggplot(running_times_longer, aes(Athlete, times, color = Month)) +  
  geom_point(group = running_times_longer$Month)
```

### We can see from the scatter plot that all runners times decreased after 3 months of the training regime.  
### Debi's & Sukhveer's time improved the most between Month 2 and 3. 
### Joana's time improved the most between Month 0 and 1.
### Kerol's time increased slighly between Month 0 and 1. Kerol's time decrease between Month 1 and 2 and Month 2 and 3 is evenly distibuted.
