---
title: "607 Project 2 - Part 1"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Practice preparing different datasets for downstream analysis work. Choose any three of the “wide” datasets identified in the Week 5 Discussion items. For each of the three chosen datasets:

(1) Create a .CSV file that includes all of the information included in the dataset. 

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

(3) Perform the analysis requested in the discussion item.

### There are two tables here with the cases each year in a separate column:

### (1) West Nile virus disease cases reported to CDC by state of residence, 1999-2019

### (2) West Nile virus neuroinvasive disease cases reported to CDC by state of residence, 1999-2019

### Pivot the year columns longer & combine data from the two tables into one. Calculate the totals, means, medians, & standard deviation per state & per year. Calculate the percent of neuroinvasive disease cases out of virus cases (per year / per state).

## Load libraries

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(tidyverse)
```

## Download .pdfs, convert to .txt, open in Excel & save as .csv files.
## Read .csv files into R:
```{r}
West_Nile_cases <- read_csv("https://raw.githubusercontent.com/vmceleney/DATA-607/main/West-Nile-virus-disease-cases-by-state_1999-2019-P.csv", 
    col_types = cols(`1999` = col_number(), 
        `2000` = col_number(), `2001` = col_number(), 
        `2002` = col_number(), `2003` = col_number(), 
        `2004` = col_number(), `2005` = col_number(), 
        `2006` = col_number(), `2007` = col_number(), 
        `2008` = col_number(), `2009` = col_number()), 
    skip = 5)
West_Nile_cases

West_Nile_neuroinvasive_cases <- read_csv("https://raw.githubusercontent.com/vmceleney/DATA-607/main/West-Nile-virus-neuroinvasive-disease-cases-by-state_1999-2019-P.csv", 
    col_types = cols(`1999` = col_number(), 
        `2000` = col_number(), `2001` = col_number(), 
        `2002` = col_number(), `2003` = col_number(), 
        `2004` = col_number(), `2005` = col_number(), 
        `2006` = col_number(), `2007` = col_number(), 
        `2008` = col_number(), `2009` = col_number()), 
    skip = 5)
West_Nile_neuroinvasive_cases
```

## Split rows for 1999 - 2009 & 2010 - 2019:
```{r}
West_Nile_cases_1999_2009 <- slice(West_Nile_cases, 1:52)
West_Nile_cases_2010_2019 <- slice(West_Nile_cases, 64:115)
West_Nile_neuroinvasive_cases_1999_2009 <- slice(West_Nile_neuroinvasive_cases, 1:52)
West_Nile_neuroinvasive_cases_2010_2019 <- slice(West_Nile_neuroinvasive_cases, 61:113)
West_Nile_cases_1999_2009
West_Nile_cases_2010_2019
West_Nile_neuroinvasive_cases_1999_2009
West_Nile_neuroinvasive_cases_2010_2019
```

## Delete empty columns:
```{r}
West_Nile_cases_2010_2019 <- select(West_Nile_cases_2010_2019, 1:11)
West_Nile_cases_2010_2019
West_Nile_neuroinvasive_cases_2010_2019 <- select(West_Nile_neuroinvasive_cases_2010_2019, 1:11)
West_Nile_neuroinvasive_cases_2010_2019
```

# Rename columns:
```{r}
West_Nile_cases_2010_2019 <- rename(West_Nile_cases_2010_2019,  
                                                  '2010' = '1999',  
                                                  '2011' = '2000',  
                                                  '2012' = '2001',  
                                                  '2013' = '2002',  
                                                  '2014' = '2003',  
                                                  '2015' = '2004',  
                                                  '2016' = '2005',  
                                                  '2017' = '2006',  
                                                  '2018' = '2007',  
                                                  '2019' = '2008')
West_Nile_cases_2010_2019

West_Nile_neuroinvasive_cases_2010_2019 <- rename(West_Nile_neuroinvasive_cases_2010_2019,  
                                                  '2010' = '1999',  
                                                  '2011' = '2000',  
                                                  '2012' = '2001',  
                                                  '2013' = '2002',  
                                                  '2014' = '2003',  
                                                  '2015' = '2004',  
                                                  '2016' = '2005',  
                                                  '2017' = '2006',  
                                                  '2018' = '2007',  
                                                  '2019' = '2008')

West_Nile_neuroinvasive_cases_2010_2019 <- slice(West_Nile_neuroinvasive_cases_2010_2019, 2:53)
West_Nile_neuroinvasive_cases_2010_2019
```

# Merge data for years 1999 - 2019
```{r}
West_Nile_cases <- inner_join(West_Nile_cases_1999_2009, West_Nile_cases_2010_2019, by = 'State')
West_Nile_cases
West_Nile_neuroinvasive_cases <- inner_join(West_Nile_neuroinvasive_cases_1999_2009, West_Nile_neuroinvasive_cases_2010_2019, by = 'State')
West_Nile_neuroinvasive_cases
```

# Pivot to longer format:
```{r}
West_Nile_cases <- pivot_longer(West_Nile_cases, cols = 2:22, names_to ="year",
values_to = "cases")
West_Nile_cases

West_Nile_neuroinvasive_cases <- pivot_longer(West_Nile_neuroinvasive_cases, cols = 2:22, names_to ="year",
values_to = "cases")
West_Nile_neuroinvasive_cases
```

# Update "Dist. of Columbia" to "Dist of Columbia' to enable join:
```{r}
West_Nile_cases$State <- str_replace_all(West_Nile_cases$State, "Dist. of Columbia", "Dist of Columbia")
```

# Merge virus case data with neuroinvasive case data:
```{r}
West_Nile <- full_join(West_Nile_cases, West_Nile_neuroinvasive_cases, by =  c('State', 'year'), suffix = c("", "_ni"))
West_Nile
```

## What is the sum, mean, median, standard deviation for each state? for each year?
```{r}
West_Nile_cases_state_summary <- West_Nile %>%
  group_by(State) %>%
  summarise(sum(cases), sum(cases_ni), mean(cases), mean(cases_ni), median(cases), median(cases_ni), sd(cases), sd(cases_ni))
West_Nile_cases_state_summary
  
West_Nile_cases_year_summary <- West_Nile %>%
  group_by(year) %>%
  summarise(sum(cases), sum(cases_ni), mean(cases), mean(cases_ni), median(cases), median(cases_ni), sd(cases), sd(cases_ni))
West_Nile_cases_year_summary
```

## Do my data totals sync up?
```{r}
sum(West_Nile_cases_state_summary$`sum(cases)`)
sum(West_Nile_cases_year_summary$`sum(cases)`)
sum(West_Nile_cases_state_summary$`sum(cases_ni)`)
sum(West_Nile_cases_year_summary$`sum(cases_ni)`)
```

## What is the proportion of neuroinvasive to virus cases for each state? for each year?
```{r}
West_Nile_cases_state_summary <- West_Nile_cases_state_summary %>%
  mutate(prop = `sum(cases_ni)` / `sum(cases)`, .after = `sum(cases_ni)`)
West_Nile_cases_state_summary
  
West_Nile_cases_year_summary <- West_Nile_cases_year_summary %>%
  mutate(prop = `sum(cases_ni)` / `sum(cases)`, .after = `sum(cases_ni)`)
West_Nile_cases_year_summary
```

### Which years had the highest proportion of neuroinvasive to virus cases?  Which years have the lowest? 
## Plot proportion of neuroinvasive per year:
```{r}
ggplot(West_Nile_cases_year_summary, aes(x= year, y = prop)) +  
  geom_point()
```

## Conclusion 
1999 - 2001 had the highest proportion of neuroinvasive to virus cases greater than 90%.  Virus detection has likely evolved since 2001 and possibly non-neuroinvasive cases were undiagnosed before 2002.