---
title: "607 TidyVerse CREATE assignment"
author: "Victoria McEleney"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

## Load libraries
```{r}
library(readr)
library(tidyverse)
```

## Bad teeth, sugar and government health spending

Source: https://www.kaggle.com/angelmm/healthteethsugar

We will focus on Bad teeth per child (12 yr, WHO) and Sugar comsumption per person (g per day, FAO) to demonstrate some basic dplyr Relational Data Mutating Join functions.

## Read .csv's
```{r}
library(readr)
badteeth <- read_csv("https://raw.githubusercontent.com/vmceleney/DATA-607/main/badteeth.csv", 
    col_types = cols(NA..1 = col_skip(), 
        NA..2 = col_skip(), NA..3 = col_skip()), n_max = 190)
View(badteeth)

sugar_consumption <- read_csv("https://raw.githubusercontent.com/vmceleney/DATA-607/main/sugar_consumption.csv", 
    col_types = cols(NA..1 = col_skip()), n_max = 258)
View(sugar_consumption)
```

## Rename columns
```{r}
colnames(badteeth) <- c("Country", "Weighted_Avg_Badteeth")
badteeth

colnames(sugar_consumption) <- c("Country", 1961:2004)
sugar_consumption
```

## Use a "Mutating Join" to join one table to columns from another, matching values with the rows that they correspond to. Each join retains a different combination of values from the tables.

# full_join(): Join data. Retain all values, all rows. 
```{r}
full_join <- full_join(sugar_consumption, badteeth, by = "Country", copy = FALSE, suffix = c(".x", ".y"), keep = FALSE, na_matched = "na")
full_join
```
Note sugar_consumption has 258 rows, badteeth has 190 rows, and full_join has 260 rows (Cote D'Ivoire and Slovak Republic did not merge correctly because of capitalization mismatch).

# left_join(): Join matching values from y to x. Includes all rows in x.
```{r}
left_join <- left_join(sugar_consumption, badteeth, by = "Country", copy = FALSE, suffix = c(".x", ".y"), keep = FALSE, na_matched = "na")
left_join
```
Note sugar_consumption has 258 rows and left_join has 258 rows.

# right_join(): Join matching values from x to y. Includes all rows in y.
```{r}
right_join <- right_join(sugar_consumption, badteeth, by = "Country", copy = FALSE, suffix = c(".x", ".y"), 2004, keep = FALSE, na_matched = "na")
right_join
```
Note badteeth has 190 rows and right_join has 190 rows.

# inner_join(): Join data. Retain only rows with matches.
```{r}
inner_join <- inner_join(sugar_consumption, badteeth, by = "Country", copy = FALSE, suffix = c(".x", ".y"), 2004, keep = FALSE, na_matched = "na")
inner_join
```
Note sugar_consumption has 258 rows, badteeth has 190 rows, and inner_join has 258 rows (Cote D'Ivoire and Slovak Republic do exist in both, but did not merge correctly because of capitalization mismatch).







